The effects of multi environmental exposures on Tuberculosis Incidence in South East Asian and African countris

### 2025-02-23 products/manuscript/Manuscript.qmd
Working on Step two of Final Project

---
title: "The effects of multi environmental exposure on Tuberculosis Incidence in Asian and African countris"
author: ""
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
editor: 
  markdown: 
    wrap: 72
---

The structure below is one possible setup for a manuscript, or a general
data analysis project (including the course project). Adjust as needed.
You don't need to have exactly these sections, but the content covering
those sections should be addressed.

@@

*Describe your methods. That should describe the data, the cleaning
processes, and the analysis approaches. You might want to provide a
shorter description here and all the details in the supplement.*

I haven't decide about data analysis methods. However, I am thinking to
perform time series analysis and multi linear regression (if suitable)
or Gene-realized linear model (we will see after conducting data
exploration). I am still in the process of reading more papers to get
more information which methods best suit this data and answer the
research questions.

@@

Sometimes you might want to show a schematic diagram/figure that was not
created with code (if you can do it with code, do it). @fig-schematic is
an example of some - completely random/unrelated - schematic that was
generated with Biorender. We store those figures in the `assets` folder.

@@

*Write code that reads in the file and cleans it so it's ready for
analysis. Since this will be fairly long code for most dataset, it might
be a good idea to have it in one or several R scripts. If that is the
case, explain here briefly what kind of cleaning/processing you do, and
provide more details and well documented code somewhere (e.g. as
supplement in a paper). All materials, including files that contain
code, should be commented well so everyone can follow along.*

@@

<!-- -->

@@

*Use a combination of text/tables/figures to explore and describe your
data. Show the most important descriptive results here. Additional ones
should go in the supplement. Even more can be in the R and Quarto files
that are part of your project.*

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the
`../../` notation. (Two dots means a folder up). You never want to
specify an **absolute** path like `C:\ahandel\myproject\results\`
because if you share this with someone, it won't work for them since
they don't have that path. You can also use the `here` R package to
create paths. See examples of that below. I generally recommend the
`here` package.

@@

@@

>>>>

---
title: "The effects of multi environmental exposures on Tuberculosis Incidence in South East Asian and African countris"
author: ""
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
editor: 
  markdown: 
    wrap: 72
---

@@

The datasets were loded one by one, then drop all unused variable before
comined those datasets using reduce() function. Countries, which have at
least one missing years from 2000 to 2020, were dropped from the study.
To make sure that the data is clean and ready to analyse, all missing
values were dropped. This study focuses on South East Asia and Africa
Regions, where the major proportion of tuberculosis cases came.
Therefore, a newsubset for South East Asia and African countries was
created.

After data cleaning, multiple exploratory data analysis was forformed
such us summary and vizualising the data. A bloxplot was created to show
Tuberculosis Incidence in each country during the time frame. Spagheti
plot was created to visualize the trend of tuberculosis incidence and
other variables over the time. Lastly, line charts showing the trend of
all variables in each country.

The exploratory data analysis showed that the data were not normally
distributed and there was no linear relationship between tuberculosis
incidence and other predictors. Therefore, Spearman's correlation matrix
was perfomed to see the correlations. To have better understanding, a
heatmap matrix plot was created to show the correlation.

The main statistical analysis in thie study is longitudinal data
analysis (eithre fixed effect or mixed effect method). To see which one
is better, Hausman Test was performed to determine which model is
better.

NOTE: I still need to discuss and need more advice about the data
analysis from Dr. Handel

@@

NOTE: I AM STILL WORKING TO FIGUR THIS ONE OUT Sometimes you might want
to show a schematic diagram/figure that was not created with code (if
you can do it with code, do it). @fig-schematic is an example of some -
completely random/unrelated - schematic that was generated with
Biorender. We store those figures in the `assets` folder.

@@

Data Import and data cleaning process have been explained in the methode
part. To see the code for data import and data cleaning , please visit
working directory ("code", "processing-code", "processingfile.qdm").
Processed and clean data are saved at a folder ("data",
"processed-data"). There are many .rds files in the folder icnlduing
file for each variables, and file after combined all variables.

@@

@@

There are some exploratory data analysis conducted and resulted table,
and charts. Table 1 shows the summary statistics including min, mean,
max, and standard deviation of all variables.

NOTE: I still have issue in saving the stiled tables. I am still working
on it. I will show it next step.

@@

The line graph ( @fig-1 ) shows the trend of all variables over the
period in each country. It can be seen that the majority of country
showed the decreases trend in tuberculosis incidence over 21 years
period, except Lesotho showing a sharp increase until the middle of the
period before dramatic decrease until the end of the period. All
countries showed the increase trend in percentage of population using
using basic and safe drinking water, basic and safe sanitation, and
clean fuel and technology for cooking, except Zimbabuwe showing the
decrease in percentage of population using basic drinking water and safe
sanitation, and Senegal showing the decrease of population using clean
fuel and technology for cooking.

```{r}

#| label: fig-1
#| fig-cap: "Trends of all variable over the period in each country"
#| echo: FALSE
knitr::include_graphics(here("code","eda-code","figures","line-graph-tbivsfactors.png"))

```

( @fig-2 ) shows spaghetti plots in each variables. We can see the
trends of each country in every variables. It can be seen that generally
most of the countries show the increase trends in water and sanitation,
clean fuel use, human development index, total greenhouses, but
fluctuated trends in mean of PM2.5 expousures.

```{r}

#| label: fig-2
#| fig-cap: "Spaghetti plots in each variables"
#| echo: FALSE
knitr::include_graphics(here("code","eda-code","figures","spageti.png"))

```

Now lets take a more detail tuberculosis incidence in each country over
the year using the following box plot. Lesotho leads the incidence
followed by Myanmar and Tanzania.

```{r}

#| label: fig-3
#| fig-cap: "Spaghetti plots in each variables"
#| echo: FALSE
knitr::include_graphics(here("code","eda-code","figures","boxplot.png"))

```

@@

As first step of statistical analysis. Using scatterplot to show the linearity between tuberculosis incidence and predictors. ( @fig-4 ) shows the scatterplots. The scatterplots do not show linear relationship between tuberculosis Incidence and predictors.

```{r}

#| label: fig-4
#| fig-cap: "Scatter plots between Tuberculosis Incidence and Predictors "
#| echo: FALSE
knitr::include_graphics(here("code","eda-code","figures","scatterplot.png"))

```
Before going further, we need to look at correlation between tuberculosis Incidence and all predictors.

Since they show nonlinearity, spearman's correlation performed and hsowing in the following correlation matrix. See ( @fig-5 )

```{r}

#| label: fig-5
#| fig-cap: "Spearman Correlation Matrix "
#| echo: FALSE
knitr::include_graphics(here("code","eda-code","figures","correlation_matrix.png"))

```

Since there is no linear relationship, nonlinear relationship test was performed.

```{r}

```
